---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.





install.packages("devtools", dependencies = TRUE)

library("devtools")

install_github("r-lib/vroom")

library(vroom)

library(tidyverse)


##read in the data
wide_spp.1 <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%203/to_sort_pop_1.csv")
wide_spp.2 <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%203/to_sort_pop_2.csv") 

## code to reshape data
## first join the data using full join - this will keep all of the columns
long_spp <- full_join(wide_spp.1, wide_spp.2) %>%
  ## pivot the joined data frame, using species, primary_threat, secondary_threat, tertiary_threat as ID columns
  ## and using names-pattern to pull out the population number 
  ## and make a new column (called population) to store them in. 
  ##Drop the NAs. 
  pivot_longer(cols = -c(species, 
                         primary_threat, 
                         secondary_threat, 
                         tertiary_threat), 
               names_to = c("population", "date"),
               names_pattern = "(.*)_(.*)",
               values_drop_na = F, 
               values_to = "abundance")

