---
title: "Week 6 Rmarkdown"
output: github_document
---
# Part 1 - Tokyo Olympics
## Is there a significant effect of a country's GDP on their position in the Olympics?
### Installing packages and data

Installing tidyverse and vroom 

```{r, echo = FALSE }
library(tidyverse)
library(vroom)
```

Install the Tokyo Medals Data:
```{r}
Tokyo_Data <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%205/Tokyo%202021%20medals.csv")
```

Install wbstats for access to the GDP data 

install.packages("wbstats")
```{r, include = FALSE}
library(wbstats)
```

Renaming the column so it is nicer to look at and easier to understand 
```{r}
my_indicators = c("GDP" = "NY.GDP.MKTP.CD")
```

Extract the GDP data for all countries
```{r}
GDP_Raw <- wb_data(my_indicators, 
                    start_date = 2020,
                    end_date = 2020)
```
Having a look at the data
```{r}
GDP_Raw
```

### Making it neater
It seems as though there are some columns we won't need, so let us get rid of those: 
Getting rid of useless columns from the data set: 
```{r}
GDP_Data <- subset(GDP_Raw, select = c(1, 2, 3, 4, 5, 9))
# Have a look
GDP_Data
# Yay it worked! 
```

## Joining the two data sets: 

Now we need to join the so let's rename one of the GDP_Data rows for the join:
```{r}
GDP_Data_Final<- rename(GDP_Data, code = iso3c)
```

Adding a column to Tokyo Data containing country code:
```{r}
library(countrycode)
```

Adding the iso3c column to the Tokyo Data enabling joining later on:
```{r}
Tokyo_Data$code <- countrycode(Tokyo_Data$Country, 
                                  origin = "country.name", 
                                  destination = "iso3c")
```

Giving China the CHN
```{r}
Tokyo_Data$code <- sub("TWN", "CHN", Tokyo_Data$code)
Tokyo_Data$code[42] <- "TWN"
```

Joining the data using inner join :
```{r}
GDP_And_Medals_Raw <- inner_join(Tokyo_Data, GDP_Data_Final, by = "code")
GDP_And_Medals_Raw
```
Looks good although we now have NA values too 

Remove all of the NA values 
```{r}
GDP_And_Medals_Final <- na.omit(GDP_And_Medals_Raw)
GDP_And_Medals_Final
```
Ordering the data set based on how the olympic medal table ranking works: 
```{r}
GDP_And_Medals_Final <-GDP_And_Medals_Final[order(-GDP_And_Medals_Final$Gold, -GDP_And_Medals_Final$Silver, -GDP_And_Medals_Final$Bronze),]
```

Using the index to give the countries a Rank:
```{r}
GDP_And_Medals_Final <- tibble::rowid_to_column(GDP_And_Medals_Final, "Rank")
```

### Getting ready to plot 
install the ggplot2 package
```{r}
install.packages("ggplot2")
# Use the ggplot 2 library 
library(ggplot2)
```

## Plotting 
Investigating an initial plot of medal ranking and GDP:

```{r GDP_And_Medals_Final, echo=FALSE}
p1 <- ggplot(GDP_And_Medals_Final, aes(x = GDP,
                                 y = Rank)) +
  geom_point() +
  theme_bw() +
  ylab("Medal Ranking") +
  xlab("GDP") 
p1
```

It looks okay but could do with some logging perhaps of the GDP to make things more clear. 

Pseudo logging the GDP and adding a geom_smooth line
```{r}
p2 <- p1 + 
  scale_x_continuous(trans="pseudo_log") + 
    geom_smooth(method="glm")
```
```{r initial plot, echo=FALSE}
p2
```

It looks okay!

## Assessing the fit of a model:
Mod 1 is going to be the "Gaussian" or normally distributed:
```{r}
mod1 <- glm(Rank ~ GDP,
                data = GDP_And_Medals_Final,
                family = "gaussian")
```

Mod 2 is going to be the "poisson" type:
```{r}
mod2 <- glm(Rank ~ GDP,
            data = GDP_And_Medals_Final,
            family = "poisson")
```
Mod 3 is the Gaussian with the link being log:
```{r}
mod3 <- glm(Rank ~ GDP,
            data = GDP_And_Medals_Final,
            family = gaussian(link="log"))
```
Mod 4 is the Gaussian link with inverse link:
```{r}
mod4 <- glm(Rank ~ GDP,
            data = GDP_And_Medals_Final,
            family = gaussian(link = "inverse"))
```
Getting library gamlr
```{r}
library(gamlr)
```

Comparing the models:
```{r}
AIC_mods <- AIC(mod1,
                mod2,
                mod3,
                mod4)
```

Rank them by AIC using the order() function
```{r}
AIC_mods[order(AIC_mods$AIC),]
```
It looks as though model 4, the gaussian, link = inverse, is slightly better so lets check that out: 

## Investigating our model further: 

Now looking at mod4
Create column of log predicted Gaussian
```{r}
GDP_And_Medals_Final$pred_gaussian_log <- predict(mod4,
                                        type = "response")
```

Plotting the diagnostics plot for model 4
```{r diagnostics, echo = FALSE}
plot(mod4)
```

Creating a column for residual Gaussian log
```{r}
GDP_And_Medals_Final$resid_gaussian_log <- resid(mod4)
```

Plotting with the geom line for the predicted Gaussian log
```{r predicted gaussian, echo = FALSE}
p3 <- ggplot(GDP_And_Medals_Final, aes(x = GDP,
                                       y = Rank)) +
  geom_point() +
  theme_bw() +
  ylab("Medal Ranking") +
  xlab("GDP") 

p3

```

Applying the model to the plot (method = glm)
```{r mod 4 and the plot}
p3 <-  p3 + 
  scale_x_continuous(trans="log") + 
  geom_smooth(method="glm") +
  xlab("Medal Ranking") +
  ylab("log_GDP") +
  ggtitle( " Logged GDP Data")

p3
```

I thought this was a good model. 
There is no residuals plotted but the data points are clear. 

Now having a look at the model summary:
```{r}
summary(mod3)
```



## This was the solution that you gave which I have adapted to my data:
```{r, echo = FALSE}
p4 <- ggplot(GDP_And_Medals_Final, aes(x=GDP, y=Rank)) + 
  geom_point() + 
  scale_y_continuous(trans='log10') + 
  scale_x_continuous(trans='log10') + 
  theme_bw() +
  ggtitle("Logged data")
  
p4
```

Fit a model where neither the x or y are logged:
```{r}
mod6 <- glm(Rank ~ GDP, data=GDP_And_Medals_Final)
plot(mod6)
```
DOESNT LOOK GOOD!

Fit a model where both the x and y are logged:
```{r}
mod7 <- glm(log10(Rank) ~ log10(GDP), data=GDP_And_Medals_Final)
plot(mod7)
```
 
So, is there an effect?
```{r}
summary(mod7)
```


# Part 2 - Iris

## Is there a positive effect of Petal.Length on Petal.Width? 

```{r}
data("iris")
library(ggplot2)
```

Plotting the sepal widths so we can visualize if there are differences between the different species
```{r, echo = FALSE}
ggplot(iris, aes(x=Petal.Length, y=Petal.Width)) + 
  geom_jitter(aes(col=Species)) +
  theme_bw()
```

Petal length and petal width histograms:

```{r}
# 1. Petal Length Histogram
a <- ggplot(iris, aes(x=Petal.Length,
                  fill=Species)) +
  geom_histogram(binwidth=.1, alpha=.5, position = "identity")
  
# 2.Petal Width Histogram
  
b <- ggplot(iris, aes(x=Petal.Width,
                 fill=Species)) +
  geom_histogram(binwidth=.1, alpha=.5, position = "identity")
```

```{r, figures-side, fig.show="hold", out.width="50%", echo = FALSE }
a
b
```


Generate and plot the model: 
```{r echo = FALSE}
mod1 <- glm(Petal.Width ~ Petal.Length * Species,
                data = iris,
                family = "gaussian")
```

Investigating the model:
```{r}
class(mod1)
plot(mod1)
summary(mod1)
```

Generating 3 other models:
```{r, echo = FALSE}
mod2 <- glm(Petal.Width ~ Petal.Length * Species,
                  data = iris,
                  family = gaussian(link = "log"))

mod3 <- glm(Petal.Width ~ Petal.Length * Species, 
                  data = iris,
                  family = "poisson")

mod4 <- glm(Petal.Width ~ Petal.Length * Species,
                  data = iris,
                  family = gaussian(link = "inverse"))
```

```{r, echo = FALSE}
library(gamlr)
```

Comparing the models using the Akaike's Information Criterion (AIC)
```{r, echo = FALSE}
AIC_mods <- data.frame(model = c("mod1", "mod2", "mod3", "mod4"),
                       AICc = c(AICc(mod1), AICc(mod2), AICc(mod3), AICc(mod4)))
```

Ranking them them by AIC using the order() function
```{r}
AIC_mods[order(AIC_mods$AICc),]
```
It looks as though model 1 is the most appropriate in this case. 

Returning the predicted gaussian and residual gaussian and adding them as columns: 
```{r}
# predicted gaussian:
iris$pred_gaussian <- predict(mod1,
                                    type="response")

# returning the residual gaussian: 
iris$resid_gaussian <- resid(mod1)
```

## Making a plot
```{r, echo = FALSE}
plot_1 <- ggplot(data = iris, aes(x = Petal.Length,
                              y = Petal.Width,
                              col = Species)) +
  geom_point() +
  geom_smooth(data = iris,
              method = glm) 


plot_1
```



## Starting Again:

```{r, echo = FALSE}
data(iris)
```

Data vizualisation: 
```{r}
iris1 <- ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) + 
  geom_point(aes(col = Species)) +
  theme_bw()

iris1
```

```{r}
iris1 + geom_smooth(aes(col = Species), method="lm")
```

Running model one 
```{r, echo = FALSE}
mod1 <- glm(Petal.Width ~ Petal.Length*Species, data = iris)
```

Set a 2x2 plot area, so we get a single pannel with 4 plots:
```{r, echo = FALSE}
par(mfrow = c(2, 2))
```

The qqplot looks a bit uneven
```{r, echo = FALSE}
plot(mod1)
```

What about square root?
```{r, echo = FALSE}
mod3 <- glm(sqrt(Petal.Width) ~ Petal.Length*Species, data = iris)
```

This looks much better:
```{r}
plot(mod3)

summary(mod3)
```










