---
title: "Week 3 Homework"
author: "Freddie Grogono"
date: "13/10/2021"
output: html_document
---

# Installing Packages
Firstly, it is important to install the relevant packages, in this case **library(tidyverse)** and **library(devtools)**.

This is a package containing functions that help with R package development.

```{r}
library("tidyverse")
library("devtools")
```

Next we install the vroom package from GitHub:
```{r, eval = FALSE}
install_github("r-lib/vroom")
```

Tell R to use the vroom() function from the vroom package: 
```{r, eval = FALSE}
vroom::vroom()
```

Make sure vroom is loaded into your R session (only need to do once):
```{r}
library(vroom)
```

Next we load the data that we are working on into the file itself.
We have already loaded the two data sets "**to_sort_pop_1.csv**" and "**to_sort_pop_2.csv**" into my GitHub repository. 

So lets first set the directory so that it knows where it is drawing the data from: 

```{r}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```


Next we identify the data within the repository and load it in:

```{r}
to_sort_pop_1 <- vroom("../Data/Workshop 3/to_sort_pop_1.csv")
to_sort_pop_1

to_sort_pop_2 <- vroom("../Data/Workshop 3/to_sort_pop_2.csv")
to_sort_pop_2
```

Using tidyverse, join both of these data sets together into a single tibble:

```{r}
Sorted_Pop_Data <- full_join(to_sort_pop_1, to_sort_pop_2)
```


Next we assign Sorted_Pop_Data values to Sorted_Pop_Data_Final AFTER performing the intended function:

```{r}
Sorted_Pop_Data_Final <- Sorted_Pop_Data %>%
# Next we pivot, 
  pivot_longer(cols = -c(species:tertiary_threat),
               names_to = c("Population","Date"),
               values_to = "Abundance",
               names_pattern = "pop_?(.*)_(.*)",
               )
Sorted_Pop_Data_Final
```

