---
title: "Week 3 Homework"
author: "Freddie Grogono"
date: "13/10/2021"
output: html_document
---

# 1. Installing Packages
## What to install?

It is important to install the relevant packages, in this case we want: **library(tidyverse)** and **library(devtools)**. 

### Tidyverse
The tidyverse is a coherent system of packages for data manipulation, exploration and visualization that share a common design philosophy <https://www.tidyverse.org/>.

### DevTools
We will also be installing devtools which is a package that helsp to simplify some tasks.

```{r, echo = TRUE}
library("tidyverse")
library("devtools")
```

### Vroom Package 

This is a package containing functions that help with R package development.
- Vroom reads rectangular data into R, such as: comma separated (csv), tab separated (tsv) and fixed width files (fwf).  

It performs similar roles to functions like: readr::read_csv (), data.table::fread (), read.csv ().

We install the vroom package from **GitHub**:
```{r, eval = TRUE}
install_github("r-lib/vroom")
```

Tell R to use the vroom() function from the vroom package: 
```{r, eval = TRUE}
vroom::vroom()
```

Make sure vroom is loaded into your R session (only need to do once):
```{r, eval = TRUE}
library(vroom)
```


# 2. Data Loading

### GitHub Repository
Next we load the data that we are working on into the file itself.
We have already loaded the two data sets "**to_sort_pop_1.csv**" and "**to_sort_pop_2.csv**" into my GitHub repository. 

So lets first set the directory so that it knows where it is drawing the data from: 

```{r, eval = FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```


Next we identify the data within the repository and load it in:

```{r}
to_sort_pop_1 <- vroom("../Data/Workshop 3/to_sort_pop_1.csv")
to_sort_pop_1

to_sort_pop_2 <- vroom("../Data/Workshop 3/to_sort_pop_2.csv")
to_sort_pop_2
```

# 3. Joinning Data
Using tidyverse, join both of these data sets together into a single tibble:

```{r}
Sorted_Pop_Data <- full_join(to_sort_pop_1, to_sort_pop_2)
# Note: the use of 'full_join' here, this is just one of many ways that you can fuse
# data together, more can be found at: <https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/join> 
```



# 4. Sorting Data

### Pivot Function
Next we need to pivot and sort the data using the pivot_longer function to 
change the data from wide to long: 

Sorted_Pop_Data values to Sorted_Pop_Data_Final AFTER performing the intended function:

```{r}
Sorted_Pop_Data_Final <- Sorted_Pop_Data %>%
  pivot_longer(cols = -c(species:tertiary_threat),
               names_to = c("Population","Date"),# This takes all the pop_1_date etc and puts them in a column called population
               values_to = "Abundance", # This takes the values of each of these dates and places them into a column called Abundance 
               names_pattern = "pop_?(.*)_(.*)",
               )
Sorted_Pop_Data_Final
```

